A Comparison of Graph Processing Systems

### 2020-09-10 documentation/paper/sections/graphFormats.tex
weitere fixes zur Rückmeldung
kleine fixes von der Rückmeldung und Polymer Overview

\section{Important Graph Formats}
Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.
Data Sets retreived from KONECT can be directly read and translated.

//SNAP

@@

In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values and \texttt{float} out-edge weights.
Each line in the graph file looks as follows
\begin{equation*}
	[s,v_s,[[t_1, w_{t_1}], [t_2, w_{t_2}]...]]
\end{equation*}
with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.
There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

>>>>

\section{Graph Formats}
Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.

The two most popular graph databases are those associated with the Koblenz Network Collection (KONECT) \cite{konect} and Stanford Network Analysis Project (SNAP) \cite{snap}.

Data Sets retreived from KONECT can be directly read and translated.

\todo{SNAP?}

@@

In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values and \texttt{float} out-edge weights.
Each line in the graph file looks as follows
\begin{equation*}
	[s,v_s,[[t_1, w_{t_1}], [t_2, w_{t_2}]...]]
\end{equation*}
with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.
%There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

### 2020-09-10 documentation/paper/sections/overview/polymer.tex
kleine fixes von der Rückmeldung und Polymer Overview

Polymer is very similar to Ligra, in fact Polymer inherits the programming interfaces \texttt{EdgeMap} and \texttt{VertexMap} from Ligra as its main interface.\cite{Polymer}

Polymer is a vertex-centric framework, that tries to circumvent some of the random memory access drawbacks of such a design. It treats a NUMA machine as a distributed cluster and splits work and graph data accordingly between the nodes.
Application-defined data is not distributed. Other runtime state data is allocated in a distributed way but only accessed through a global lookup table.

>>>>

Polymer is a NUMA-aware graph-analytics system that inherits the scatter- gather programming interface from Ligra \cite{Ligra}. 
Key differences are the data layout and access strategies, Polymer implements.
The goal is to minimize random and remote memory accesses to improve performance. 
%Computation is performed in a vertex-centric manner.

The first optimization Polymer applies is data locality and access methods across NUMA nodes \cite{Polymer}.
A general design principle for NUMA machines is to partition the input data so that computation can be grouped with the corresponding data on one node.
Polymer adopts this and allocates graph data according to the access patterns.
It treats a NUMA machine like a distributed cluster and splits work and graph data accordingly between the nodes. 
Vertices and Edges are partitioned and then allocated across the corresponding memory nodes of the threads, eliminating most remote memory accesses.
However, some computation requires vertices to perform computations on edges that are not in the local NUMA-node.
For this case, Polymer introduces lightweight vertex replicas that are used to initiate computation on remote edges.

Polymer furthermore has custom storage principles for application-defined data.
For such data, the memory locations are static but the data undergo frequent dynamic updates. Due to frequent exchanges of application-defined data between the nodes, remote memory accesses are inevitable.
Hence, Polymer allocates application-defined data with virtual adresses, while distributing the actual memory locations across the nodes. Thus, all updates are applied on a single copy of application-defined data.

Runtime states that are dynamically allocated in each iteration would however create overhead due to repeated construction of a virtual address space.
These states are thus stored in a custom lock-less (i.e. avoiding contention) lookup table.

Polymer does not only optimize data locality but also its scheduling is custom.
Time to synchronize threads on different cores increases dramatically with the growing number of involved sockets. Inter-node synchronization takes one order of magnitude longer time than intra-node synchronization \cite{Polymer}.
Thus, Polymer implements a topology-aware hierarchical synchronization barrier.
A group of threads on the same NUMA-node shares a partition of data. This allows them to first only synchronize with threads on the same NUMA-node. Only the last thread of each group sychronizes across groups (i.e. nodes). This behaviour decreases the amount of needed cache coherence broadcasts across the nodes.

Furthermore, Polymer switches between different data structures representing the runtime state, similiar to Ligra's behaviour. The main deciding factor to switch is the amount of active vertices relative to an application-defined threshold. Polymer uses a lock-less tree structure representing the active vertices. The leaves use bitmaps, which are efficient for a large proportion of active vertices. When only a small amount of vertices is active, the drawbacks of traversing through sparse bitmaps can be avoided by switching data structures.

Polymer inherits the programming interfaces \texttt{EdgeMap} and \texttt{VertexMap} from Ligra as its main interface.

### 2020-09-10 documentation/paper/sections/preliminiaries.tex
kleine fixes von der Rückmeldung und Polymer Overview

\subsection{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set is defined as 
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E$, it holds $(x,y)=(y,x)$. 
The size of a graph is defined as the number of edges $|E|$\todo{citation?}.
Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case a function $w:E\rightarrow \mathbb R$ is introduced, that maps an edge to a numerical value, further describing the relation.

@@

\subsection{Single-Source Shortest-Paths}
Single-Source Shortest-Paths (SSSP) describes the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.
The most common sequential implementations are Dijkstra's algorithm or BellmanFord \cite{Polymer, Ligra, pregel}.

@@

\subsection{PageRank}
\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}. This represents a centrality metric of the vertices.
The analogy is that the graph represents website pages of the Word Wide Web, that are hyperlinked between one another. A website that is more important is likely to receive more links from other webiste.
PageRank counts the number and quality of links to a page to estimate the importance of a page.
The output of PageRank is a percentage for each vertex. This percentage, called the PageRank of the vertex, is the probability with which a web surfer starting at a random web page reaches this webpage (vertex). 
With a high probability the web surfer uses a random link from the web page they are currently on and with a smaller probability (called damping factor) they jump to a completely random web page.

The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.

>>>>

\subsection{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set is defined as 
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E$, it holds $(x,y)=(y,x)$. 
The size of a graph is defined as the number of edges $|E|$ \cite{newman2010networks}.
Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case a function $w:E\rightarrow \mathbb R$ is introduced, that maps an edge to a numerical value, further describing the relation.

@@

\subsection{Single-Source Shortest-Paths}
Single-Source Shortest-Paths (SSSP) describes the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
Input to the problem is a weighted graph $G=(V,E)$ and a start vertex $s\in V$. Output is the shortest possible distance from $s$ to each vertex in $V$. 
The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.
The most common sequential implementations are Dijkstra's algorithm or BellmanFord \cite{Polymer, Ligra, pregel}.

@@

\subsection{PageRank}
\todo{Aufbau unschön, Formel fehlt}
\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}. 
The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.

This represents a centrality metric of the vertices.
The analogy is that the graph represents website pages of the Word Wide Web, that are hyperlinked between one another. A website that is more important is likely to receive more links from other webiste.
PageRank counts the number and quality of links to a page to estimate the importance of a page.
The output of PageRank is a percentage for each vertex. This percentage, called the PageRank of the vertex, is the probability with which a web surfer starting at a random web page reaches this webpage (vertex). 
With a high probability the web surfer uses a random link from the web page they are currently on and with a smaller probability (called damping factor) they jump to a completely random web page.

### 2020-09-10 documentation/paper/sections/testingMethods.tex
results for pr, removed two algos
kleine fixes von der Rückmeldung und Polymer Overview

\subsection{Testing Methods}
For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each. One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
All five machines were running Ubuntu 18.04.2 LTS.

Setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
All benchmark cases were initiated by our benchmark script available in our repository. 
All five frameworks are tested on a single server. 
Galois, Gemini and Giraph were benchmarked in on the distributed 5-node cluster as well.
Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
Unless mentioned otherwise, we always show results of each framework utilizing 96 threads (i.e. the maximum on our machines) for the single-node evaluation.

>>>>

\subsection{Environment}
For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each. One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
All five machines were running Ubuntu 18.04.2 LTS.

Setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
All benchmark cases were initiated by our benchmark script available in our repository. 
All five frameworks are tested on a single server. 
Galois, Gemini and Giraph were benchmarked in on the distributed 5-node cluster as well.
Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
Furthermore, Galois is a framework capable of utilizing hugepages. We include an evaluation using those on the single node as well.
Unless mentioned otherwise, we always show results of each framework utilizing 96 threads (i.e. the maximum on our machines) for the single-node evaluation.
